---
title: 'Coursera: Learn Practical Machine Learning Lab Sandbox'
author: "Nisa Nuraulia"
date: "2/13/2025"
output: html_document
---

# Introduction
This project aims to predict how participants performed barbell lifts using sensor data from their **belt, forearm, arm, and dumbbell**.  
We will use **Machine Learning** to classify the exercises into five categories (**A, B, C, D, E**) based on the `classe` variable.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(rstudioapi)
library(caret)
library(randomForest)
```

```{r load-data}
# Load datasets
training_data <- read.csv("pml-training.csv", na.strings = c("NA", "", "#DIV/0!"))
test_data <- read.csv("pml-testing.csv", na.strings = c("NA", "", "#DIV/0!"))

# View structure of training data
str(training_data)
```

```{r clean-data}
# Clean dataset: Remove columns with NA values and unnecessary columns
training_data <- training_data[, colSums(is.na(training_data)) == 0]
training_data <- training_data[, -(1:7)]
training_data$classe <- as.factor(training_data$classe)
```

```{r data-splitting}
# Split data into training and validation sets
set.seed(123)
inTrain <- createDataPartition(training_data$classe, p = 0.7, list = FALSE)
train_set <- training_data[inTrain, ]
validation_set <- training_data[-inTrain, ]

# Check dimensions of training and validation sets
dim(train_set)
dim(validation_set)
```

```{r model-training}
# Train Random Forest model
set.seed(123)
rf_model <- randomForest(classe ~ ., data = train_set, importance = TRUE)
```

```{r model-evaluation}
# Make predictions on validation set
rf_predictions <- predict(rf_model, validation_set)

# Evaluate model performance
conf_matrix <- confusionMatrix(rf_predictions, validation_set$classe)
conf_matrix
```

```{r final-predictions}
# Make final predictions on test data
final_predictions <- predict(rf_model, test_data)
answers <- data.frame(problem_id = 1:length(final_predictions), classe = final_predictions)
answers
